Fault-tolerant Asynchronous Circuits

### 2019-12-04 draft_final.tex
last fair traces version
edited ch 4, changed L3

\title{Background for transient faults in asynchronous circuits}
\author{Zoey Zhou}
\date{\today}  
\usepackage{graphicx}
\graphicspath{ {c:/Desktop/} }

\begin{document}
\chapter{Introduction}

@@

\section{Synchronous vs Asynchronous circuits}
Digital circuits can be divided into two types, synchronous and asynchronous.  The time it takes for gates to process input signals and then propagate the output to the next gates in a digital circuit may vary widely.  If the timing is somehow off, it is possible for the outputs to be different from the desired value resulting in incorrect computation. In synchronous systems this timing variation is dealt with by using a global clock.  The clock is then fed into each flip flop so that the flip flop outputs will lock and update the signals synchronously. The clock cycle frequency is chosen so that all of the gates can finish processing and the correct values are locked for the next cycle. That is, the clock cycle is usually designed to accommodate the slowest logic element. The design of synchronous systems can be complex and is not necessary to understand in detail for this work. \\
\begin{figure}
\includegraphics[width=\textwidth]{syncex}
\captionof{figure}{Example of a synchronous digital circuit.  Each output is gated by a flip flop that is controlled by the clock}
\end{figure}
%possible picture of clock and flip flops?.... redraw as my own

@@

\includegraphics[width=\textwidth]{}
\captionof{figure}{Example of a simple asynchronous circuit.  Note that there are no clocks and the signals can propagate at any time}

@@

%C-elements
The C-element proposed by Muller is an important building block for asynchronous circuits. It acts as a memory element in that it receives two inputs. When both inputs are low, the C-element output goes low, and when both inputs are high the C-element output goes high. However, when the inputs are mismatched then the output gets held at its previous value. This can be achieved in a number of ways such as a gate implementation as well as a transistor level implementation with feedback using a weak transistor, other implementations use capacitance to achieve this weak feedback.
\includegraphics[width=\textwidth]{celement}
\captionof{figure}{a. logic of the C-element. A, B are inputs, C is the output \\ b. diagram of a C-element at the gate level c. transistor-level implementation}

@@

Even though the generalized C-element is presented as a C-element with set and reset as inputs, the transistor level implementation differs slightly. We assume a transistor level implementation of the set and reset combined with the C-element so that there is only one delay for the entire logic block. In addition, this is not a normal C-element because set is used exclusively for the pull-up logic and reset is used for the pull-down logic. In a normal C-element we allow for any four combinations of the values of the set and reset input wires. However, using the pull-up and pull-down logic setup, we do not allow for when both set and reset is `ON' as the output wire would simultaneously be connected to the high voltage and ground. Such an output would have indeterminable voltage values and violate the digital assumptions. Thus, the set and reset logic is designed so that they cannot overlap. This corresponds to $\overline{set}\vee\overline{reset} $ is always TRUE.\\
\begin{center}
\includegraphics[width=.7\textwidth]{genC}
\end{center}
\captionof{figure}{transistor level implementation of a generalized C-element. The + signs are the set logic inputs and - signs are the reset logic inputs. For this particular element, set is in3 AND in4 while reset is in1 AND in2.}

@@

Triple Modular Redundancy is a technique where each combinational logic unit is copied three times and then fed into a latch that performs a majority voting operation. That is, if at least 2 out of the 3 signals agree on a value, the output becomes that value. Additionally, to counter possible faults on the single output wire, the majority vote unit can also be copied three times and its output wire fed appropriately as input to the three copies of the downstream combinational logic. This would resolve both transient and permanent single faults. However, the downside is that the circuit needs to be copied three times so the area is three times larger and the power consumption as well. In addition, there must be many adaptations for this to work in asynchronous circuits as the signals can arrive in different orders. We give an example of why TMR does not work in an asynchronous circuit as it does in a synchronous circuit below. We assume that in the original circuit that each logic block has some input wires and an output wire. In the TMR scheme, each logic block is triplicated and to remove single points of failure, we also triplicate the majority voter that the three copies of the output wire feeds into. There are also three distinct copies of each input signal that correspond to outputs from the three voters. This is the basic scheme that synchronous circuits with TMR uses. Because this circuit is asynchronous, we assume that each logic block is composed of gates that has some arbitrary delay that is independent of other gates. This is true under both delay assumptions of speed independent and delay insensitive circuits. We ignore whether a fault can be in the voter for now. 
\begin{center}
\includegraphics[width=\textwidth]{tmrcounterex}
\end{center}
\captionof{figure}{An example of an error if a fault occurs in an asynchronous TMR circuit}
Initially all output wire values are `0's. Then since the input signals are distinct copies of each other, it is possible for one set to finish transitions (new) before the other two sets and the first logic block transitions to a `1'. Then a transient fault arrives on the output of the second logic block which transitions to a `1' as well. The majority voter would then transition to a `1'. However, the upset from transient faults usually happen within a short time period and the old inputs eventually force the output of the second logic block back to `0'. Then the majority voter would transition to a `0' so that causes a hazard on the output of the majority voter. %depends on if the logic gate was combinational logic.... which I guess we can assume it is
%NEEDS FIXING AND CHECKING
It is also possible for the `1' from the majority voter to feed into other logic units which is effectively an early transition. That early transition will then put the circuit into states that are not specified by the original state graph and thus the next states cannot be predicted.
\\

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{gates1}
\captionof{figure}{a) the AND element is excited  b) an excited element fires to become stable}
\end{figure}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{element1}
\captionof{figure}{a) the C-element is excited  b) two types of stable C-elements}
\end{figure}

@@

 % A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.
When we specify a circuit, we must also specify the initial state.  
There are some additional properties that a trace can have and we define them here.
\begin{definition}
A {\em stopping state} is a state where there are no excited elements in the circuit. 
\end{definition} 
 The circuit may be waiting for a primary input change.
%Thus, the only possible transitions from a stopping state are from changes in the primary inputs.  In other words, in a stopping state, the circuit is waiting for the primary inputs to changes.\\
\begin{definition}
A {\em complete trace} is either a finite length trace that ends in a stopping state, or an infinite trace.
\end{definition}
A complete trace of finite length must end in a stopping state where the primary inputs never change and the circuit is stopped.  In other words, for any state of a complete trace, if there are excited elements, one of them must fire, causing the circuit to transition to another state.
\begin{definition}
A {\em fair trace} is a trace of a circuit where if an element is excited in any state, it becomes stable in a finite number of transitions.
\end{definition}
In other words, every circuit element must be stable infinitely often in a fair trace.  This fits with the physical properties of circuits because the elements are not arbitrarily slow and if excited, they will eventually change. 
A set of traces can describe a circuit completely, in addition we may be interested in the states that the circuit visits in these traces. 
A {\em reachable state} is defined in this way.
\begin{definition}
A state $s$ is {\em reachable} from $s_1$ if $s$ appears in some trace that begins with $s_1$.
%A {\em reachable state} is an element of the set of all states in some trace, of a set of traces. 
%still kind of weird oh well 
\end{definition}
Thus, unreachable states are states that are not in any of the traces with initial state $s_1$.  Since all functionality of a circuit can be described by a set of traces starting in a specified state, then the circuit will never enter those unreachable states.

@@

Combining semi-modularity and fair traces, since excited elements in semi-modular circuits can only become stable from the element firing, then a fair infinite trace of the circuit has the property that if an element is excited infinitely often, then it fires infinitely often.

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\chaptermark{Fault Tolerant Circuit}
\section{Traces and circuit equivalence}
We solve the problem of single transient faults by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all of its circuit elements.  Due to the asynchronous nature of how signals are processed, corresponding wires in the resulting half circuits may have different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.   An example of this transformation applied to the simple inverter circuit is shown in figure \ref{fig:dupschemeex}.  Both the duplication and the additional C-elements configuration are shown.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault-tolerant circuit is equivalent to the original.  Finally we show that if a single transient fault occurs, one side of the fault-tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\
\begin{figure}
  \centering
    \includegraphics[width=.9\textwidth]{transf3Csimp}
  \caption{a) original circuit b) after the fault-tolerant transformation}
  \label{fig:dupschemeex}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=.6\textwidth]{sm_counter2}
  \caption{a) original circuit for wire w b) an erroneous initialization}
  \label{fig:sm_counter2}
\end{figure}

@@

We can make an existing semi-modular circuit fault-tolerant by the following transformation as shown in figure \ref{fig:dupscheme}: we make a duplicate copy of the logic blocks of each wire $w$ and label the duplicate wires $w^A$ and $w^B$.  We also label the duplicated logic blocks as belonging to circuit A or circuit B in agreement to the wire naming.  Then we connect the output of the logic blocks to three layers of C-elements as shown in figure \ref{fig:dupscheme}.  We label the intermediate wires as $c1^A$, $c2^A$, $c3^A$ for circuit A and $c1^B$, $c2^B$, $c3^B$ for circuit B.  In the first layer of C-elements, $c1^A$ and $c1^B$ are outputs of the logic blocks and inputs to two C-elements outputting $c2^A$ and $c2^B$.  These are then the inputs to the second layer of C-elements outputting $c3^A$ and $c3^B$.  These are then the inputs to the third layer of C-elements outputting $w^A$ and $w^B$.  We can label one of the final C-element outputs as $w^A$ and the other output as $w^B$.  These then connect to other logic blocks in circuit A and circuit B respectively.  \\
Note that when labelling the C-elements and its outputs, which side is A and B can be somewhat arbitrary because the output pairs are symmetric.  The symmetry is broken when connecting the final output wire $w^A$ to all the circuit A gate inputs and wire $w^B$ to all the circuit B gate inputs.  All circuit elements with the A label is part of the A-half circuit and all circuit elements with the B label is part of the B-half circuit.\\%(Note that the outputs of the C-elements can be arbitrarily named and connected to the next gates, and the next gates can also be arbitrarily named $A$ or $B$ as long as it does not create any inconsistencies in the naming convention, ie. all $A$ wires feed into the $A$ gates)  %the only wires that really matter are the wires that have isochronic forks, these must input to same A or B, everything else is just a naming difference
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{circuitforproof3}
  \caption{Fault-tolerant transformation: duplicate the original logic for each output wire and adding 3 layers of C-elements}
  \label{fig:dupscheme}
\end{figure}

@@

%[Why is this the best we can do?]
We compare each half circuit because it is not easy to synchronize and compare the signals between both halves, especially if we reuse the logic blocks from the original circuit.  A simple idea we tried was to insert another C-element before the input is fed into the logic, to synchronize the wires on both halves.  The problem with that is in the original circuit, a wire value change propagates to all inputs at the same time.  Adding another layer of C-elements creates delays at the input to logic blocks and the isochronic fork assumption is violated.  Thus the resulting circuit will not function according to the desired specifications.  (an example can be provided but not sure if needed?)\\  
\begin{figure}
  \centering
    \includegraphics[width=.8\textwidth]{counterexhalf}
  \caption{State graph example to show how b can transition before a or c finishes in both halves in the fault-tolerant circuit}
  \label{fig:counterexhalf}
\end{figure}

@@

\begin{lemma}[0]
Given two identical logic blocks $f$ and $g$, if the inputs into $f$ and $g$ have the same values then:\\
i)  if the outputs are the same value and $f$ is excited then $g$ is also excited.\\
ii)  if the outputs are the same value and $f$ is unexcited then $g$ is also unexcited.\\
iii)  if the outputs are different values and $f$ is excited then $g$ is unexcited.
\end{lemma}
\begin{proof}
i) \& ii)  Since $f$ and $g$ are identical gates or C-elements then if the inputs and output are the same values, $f$ and $g$ must be both excited or both unexcited. \\

iii)  We prove this by contradiction.  Suppose that inputs to $f$ and $g$ are the same and the outputs are different values and both $f$ and $g$ are excited.  This means that for some circuit element that these input values both cause the element to excite to 1 and 0 at the same time.  This cannot happen in any of the elements that we build our circuit from and is therefore a contradiction.
\end{proof}

@@

%P3 states that the wire block must be in a state a-d) of figure \ref{fig:l1helper}.  We consider what happens if a wire transitions from each state. 
%logic gate excited = c1^A excited
%possibly add graph to illustrate this
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{gatewl2}
  \caption{Different cases of P2 for a wireblock, 1-4 correspond to excited wire in the original circuit and 5 is unexcited.  1) case 1, 2) case 2, 3) case 3, 4) case 4, 5) case 5}
  \label{fig:l1helper}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics{flowl2c3}
  \caption{The wires in each row are in order of logic block, C1, C2, C3, W.  Each row represents half of the duplicated circuit, ie. top row are A half wires and bottom row are B half wires.  'Un' denotes that the logic block is unexcited and 'Exc' denotes the logic block is excited.  Blue arrows indicate a case change caused by transitions elsewhere in the circuit.  Black arrows indicate a wire transition.}
  \label{fig:l21}
\end{figure}
%fixed 
\begin{lemma}[2.1]
For any trace of the fault-tolerant circuit, each wire block follow the state diagram in Figure \ref{fig:l21} (up to 0 and 1 inversion).
\end{lemma}
\begin{proof}
We show this lemma by induction.  The state diagram in Figure \ref{fig:l21} is incomplete in the sense that after case 11, there are more transitions.  However, case 11 is symmetric to case 1 with only the 0's and 1's inverted and so the transitions from case 11 is also symmetric to the transitions from case 1 with 0's and 1's inverted until it loops back to case 1.  Because of this symmetry, we only need to analyze possible transitions from each case up to case 10.  Also, Lemma 1 applies to both A half and B half which we will use in the body of this proof.\\

@@

%still a slight problem here
\end{itemize}
This shows that all wire blocks in all possible $s_{k+1}$ follow the state diagram, and the induction step is complete.\\
By induction on the consecutive states of the trace, the above argument shows that all wire blocks of a fault-tolerant circuit follow the state diagram in Figure \ref{fig:l21} for all traces. \\
\end{proof}
%may be that I don't need how all the intermediate wires go... but just need the at most one of input excitation and outputs can be mismatched.  But this needs to include if outputs are mismatched then the C-element is the same.... which I think I need to show the signals propagating through??
It is now easy to show that deadlock does not occur since each wire block must be in one of the configurations of figure \ref{fig:l21}.
\begin{lemma}[2.2]
For all traces $\sigma$ in the fault-tolerant circuit, deadlock does not occur.
\end{lemma}
\begin{proof}
To show that deadlock does not occur, we show that we can always find an excited wire at any state $s_k$ in a trace of the fault-tolerant circuit.  The state $s_k$ can be split into two cases, one is that all output wire pairs in the A half and B half are equal, the second is that there exists at least one wire $v$ such that $v^A\neq v^B$.  In the first case, if every output wire pair are equal, then $h^A(s_n)=h^B(s_n)$ and there exists an excited wire $u$ in the original circuit.  By Lemma 1 both halves of the duplicated circuit for $u$ is excited.  By Lemma 2.1, the circuit must be in case 2-6 for $u$.  In each of these cases at least one wire is excited.  \\
In the second case, for some wire $v$ such that $v^A\neq v^B$, then using Lemma 2.1 wire block $v$ is in case 7a or 7b which has an excited wire, either $v^A$ or $v^B$.  Since there are next transitions for both cases, deadlock does not occur.
\end{proof}

>>>>

\usepackage{graphicx}
\graphicspath{ {c:/Desktop/} }

    \title{Fault-tolerant Asynchronous Circuits}
    \author{Zoey Zhou}
    \dept{Electrical Engineering}
    \principaladviser{David Dill}
    \firstreader{Subhasish Mitra}
    \secondreader{Clark Barrett\\(Computer Science)}

%\title{Background for transient faults in asynchronous circuits}
%\author{Zoey Zhou}
%\date{\today}

\begin{document}

\beforepreface

\prefacesection{Abstract}

\newpage

%% one can also have a prefacesection that is a Preface instead of
%% Acknowledgements.   The thematic purpose is the same (thanks).
    \prefacesection{Acknowledgements}
        I would like to thank Professor David Dill.

%% afterpreface produces a table of contents and any other tables
%% wanted. At the end pagenumbering changes from roman to arabic and
%% is restarted
    \afterpreface

\chapter{Introduction}

@@

\section{Synchronous vs Asynchronous circuits}
Digital circuits can be divided into two types, synchronous and asynchronous.  In digital circuits, the time it takes for gates to process input signals into output signals and for these signals to reach the next gates vary widely.  If the timing is somehow off, it is possible for the outputs to be different from the desired value resulting in incorrect computation. In synchronous systems this timing variation is dealt with by using a global clock.  The clock is then fed into each flip flop so that the flip flop outputs will lock and update the signals synchronously. The clock cycle frequency is chosen so that all of the gates can finish processing and the correct values are locked for the next cycle. That is, the clock cycle is usually designed to accommodate the slowest logic element. The design of synchronous systems can be complex and is not necessary to understand in detail for this work. \\
\begin{figure}
\includegraphics[width=\textwidth]{syncex}
\captionof{figure}[Example of a synchronous digital circuit]{Example of a synchronous digital circuit.  Each output is gated by a flip flop that is controlled by the clock}
\end{figure}
%possible picture of clock and flip flops?.... redraw as my own

@@

%\includegraphics[width=\textwidth]{}
%\captionof{figure}{Example of a simple asynchronous circuit.  Note that there are no clocks and the signals can propagate at any time}

@@

%C-elements
The C-element proposed by Muller is an important building block for asynchronous circuits. It acts as a memory element in that it receives two inputs. When both inputs are low, the C-element output goes low, and when both inputs are high the C-element output goes high. However, when the inputs are mismatched then the output gets held at its previous value. This can be achieved in a number of ways such as a gate implementation as well as a transistor level implementation with feedback using a weak transistor, other implementations use capacitance to achieve this weak feedback.
\begin{figure}
\includegraphics[width=\textwidth]{celement}
\captionof{figure}[C-element: logic and implementation]{a. logic of the C-element. A, B are inputs, C is the output \\ b. diagram of a C-element at the gate level c. transistor-level implementation}
\end{figure}

@@

Even though the generalized C-element is presented as a C-element with set and reset as inputs, the transistor level implementation differs slightly. We assume a transistor level implementation of the set and reset combined with the C-element so that there is only one delay for the entire logic block. In addition, this is not a normal C-element because set is used exclusively for the pull-up logic and reset is used for the pull-down logic. In a normal C-element we allow for any four combinations of the values of the set and reset input wires. However, using the pull-up and pull-down logic setup, we do not allow for when both set and reset is `ON' as the output wire would simultaneously be connected to the high voltage and ground. Such an output would have indeterminable voltage values and violate the digital assumptions. Thus, the set and reset logic is designed so that they cannot overlap. This corresponds to $\overline{set}\vee\overline{reset} $ is always TRUE.\\
\begin{center}
\includegraphics[width=.7\textwidth]{genC}
\end{center}
\captionof{figure}[Generalized C-element implementation]{Transistor level implementation of a generalized C-element. The + signs are the set logic inputs and - signs are the reset logic inputs. For this particular element, set is in3 AND in4 while reset is in1 AND in2.}

@@

Triple Modular Redundancy is a technique where each combinational logic unit is copied three times and then fed into a latch that performs a majority voting operation. That is, if at least 2 out of the 3 signals agree on a value, the output becomes that value. Additionally, to counter possible faults on the single output wire, the majority vote unit can also be copied three times and its output wire fed appropriately as input to the three copies of the downstream combinational logic. This would resolve both transient and permanent single faults. However, the downside is that the circuit needs to be copied three times so the area is three times larger and the power consumption as well. In addition, there must be many adaptations for this to work in asynchronous circuits as the signals can arrive in different orders. We give an example of why TMR does not work in an asynchronous circuit as it does in a synchronous circuit below. We assume that in the original circuit that each logic block has some input wires and an output wire. In the TMR scheme, each logic block is triplicated and to remove single points of failure, we also triplicate the majority voter that the three copies of the output wire feeds into. There are also three distinct copies of each input signal that correspond to outputs from the three voters. This is the basic scheme that synchronous circuits with TMR uses. Because this circuit is asynchronous, we assume that each logic block is composed of gates that has some arbitrary delay that is independent of other gates. This is true under both delay assumptions of speed independent and delay insensitive circuits. We ignore whether a fault can be in the voter for now. 
\begin{center}
\includegraphics[width=\textwidth]{tmrcounterex}
\end{center}
\captionof{figure}{TMR asynchronous circuit example with fault}
Initially all output wire values are `0's. Then since the input signals are distinct copies of each other, it is possible for one set to finish transitions (new) before the other two sets and the first logic block transitions to a `1'. Then a transient fault arrives on the output of the second logic block which transitions to a `1' as well. The majority voter would then transition to a `1'. However, the upset from transient faults usually happen within a short time period and the old inputs eventually force the output of the second logic block back to `0'. Then the majority voter would transition to a `0' so that causes a hazard on the output of the majority voter. %depends on if the logic gate was combinational logic.... which I guess we can assume it is
%NEEDS FIXING AND CHECKING
It is also possible for the `1' from the majority voter to feed into other logic units which is effectively an early transition. That early transition will then put the circuit into states that are not specified by the original state graph and thus the next states cannot be predicted.
\\

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{gates1}
\captionof{figure}[AND excited and stable]{a) the AND element is excited  b) an excited element fires to become stable}
\end{figure}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{element1}
\captionof{figure}[C-element excited and stable]{a) the C-element is excited  b) two types of stable C-elements}
\end{figure}

@@

 % A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.
When we specify a circuit, we must also specify the initial state.  
There are some additional properties that a trace can have and we define them here.
\begin{definition}
A {\em stopping state} is a state where there are no excited elements in the circuit. 
\end{definition} 
 The circuit may be waiting for a primary input change.
%Thus, the only possible transitions from a stopping state are from changes in the primary inputs.  In other words, in a stopping state, the circuit is waiting for the primary inputs to changes.\\
\begin{definition}
A {\em complete trace} is either a finite length trace that ends in a stopping state, or an infinite trace.
\end{definition}
A complete trace of finite length must end in a stopping state where the primary inputs never change and the circuit is stopped.  In other words, for any state of a complete trace, if there are excited elements, one of them must fire, causing the circuit to transition to another state.
%\begin{definition}
%A {\em fair trace} is a trace of a circuit where if an element is excited in any state, it becomes stable in a finite number of transitions.
%\end{definition}
I%n other words, every circuit element must be stable infinitely often in a fair trace.  This fits with the physical properties of circuits because the elements are not arbitrarily slow and if excited, they will eventually change. 
A set of traces can describe a circuit completely, in addition we may be interested in the states that the circuit visits in these traces. 
A {\em reachable state} is defined in this way.
\begin{definition}
A state $s$ is {\em reachable} from $s_1$ if $s$ appears in some trace that begins with $s_1$.
%A {\em reachable state} is an element of the set of all states in some trace, of a set of traces. 
%still kind of weird oh well 
\end{definition}
Thus, unreachable states are states that are not in any of the traces with initial state $s_1$.  Since all functionality of a circuit can be described by a set of traces starting in a specified state, then the circuit will never enter those unreachable states.

@@

%Combining semi-modularity and fair traces, since excited elements in semi-modular circuits can only become stable from the element firing, then a fair infinite trace of the circuit has the property that if an element is excited infinitely often, then it fires infinitely often.

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\chaptermark{Fault Tolerant Circuit}
\section{Traces and circuit equivalence}
We solve the problem of single transient faults by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all of its circuit elements.  Due to the asynchronous nature of how signals are processed, corresponding wires in the resulting half circuits may have different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.   An example of this transformation applied to the simple inverter circuit is shown in figure \ref{fig:dupschemeex}.  Both the duplication and the additional C-elements configuration are shown.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault-tolerant circuit is equivalent to the original.  Finally we show that if a single transient fault occurs, one side of the fault-tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\
\begin{figure}
  \centering
    \includegraphics[width=.9\textwidth]{transf3Csimp}
  \caption[Fault-tolerant transformation of an example circuit]{Fault-tolerant transformation of an example circuit: a) original circuit b) after the fault-tolerant transformation}
  \label{fig:dupschemeex}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics[width=.6\textwidth]{sm_counter2}
  \caption[Initializations]{Initializations: a) original circuit for wire w b) an erroneous initialization}
  \label{fig:sm_counter2}
\end{figure}

@@

We can make an existing semi-modular circuit fault-tolerant by the following transformation as shown in figure \ref{fig:dupscheme}: we make a duplicate copy of the logic blocks of each wire $w$ and label the duplicate wires $w^A$ and $w^B$.  We also label the duplicated logic blocks as belonging to circuit A or circuit B in agreement to the wire naming.  Then we connect the output of the logic blocks to three layers of C-elements as shown in figure \ref{fig:dupscheme}.  We label the intermediate wires as $c1^A$, $c2^A$, $c3^A$ for circuit A and $c1^B$, $c2^B$, $c3^B$ for circuit B.  In the first layer of C-elements, $c1^A$ and $c1^B$ are outputs of the logic blocks and inputs to two C-elements outputting $c2^A$ and $c2^B$.  These are then the inputs to the second layer of C-elements outputting $c3^A$ and $c3^B$.  These are then the inputs to the third layer of C-elements outputting $w^A$ and $w^B$.  We can label one of the final C-element outputs as $w^A$ and the other output as $w^B$.  These then connect to other logic blocks in circuit A and circuit B respectively.  \\
Note that when labelling the C-elements and its outputs, which side is A and B can be somewhat arbitrary because the output pairs are symmetric.  The symmetry is broken when connecting the final output wire $w^A$ to all the circuit A gate inputs and wire $w^B$ to all the circuit B gate inputs.  All circuit elements with the A label is part of the A-half circuit and all circuit elements with the B label is part of the B-half circuit.\\%(Note that the outputs of the C-elements can be arbitrarily named and connected to the next gates, and the next gates can also be arbitrarily named $A$ or $B$ as long as it does not create any inconsistencies in the naming convention, ie. all $A$ wires feed into the $A$ gates)  %the only wires that really matter are the wires that have isochronic forks, these must input to same A or B, everything else is just a naming difference
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{circuitforproof3}
  \caption[Fault-tolerant transformation of each wire]{Fault-tolerant transformation: duplicate the original logic for each output wire and adding 3 layers of C-elements}
  \label{fig:dupscheme}
\end{figure}

@@

%[Why is this the best we can do?]
We compare each half circuit because it is not easy to synchronize and compare the signals between both halves, especially if we reuse the logic blocks from the original circuit.  A simple idea we tried was to insert another C-element before the input is fed into the logic, to synchronize the wires on both halves.  The problem with that is in the original circuit, a wire value change propagates to all inputs at the same time.  Adding another layer of C-elements creates delays at the input to logic blocks and the isochronic fork assumption is violated.  Thus the resulting circuit will not function according to the desired specifications.  (an example can be provided but not sure if needed?)\\  
\begin{figure}
  \centering
    \includegraphics[width=.8\textwidth]{counterexhalf}
  \caption[State graph example of fault-tolerant circuit]{State graph example to show how b can transition before a or c finishes in both halves in the fault-tolerant circuit}
  \label{fig:counterexhalf}
\end{figure}

@@

\begin{lemma}[0]
Given two identical logic blocks $f$ and $g$, if the inputs into $f$ and $g$ have the same values then:\\
i)  if the outputs are the same value and $f$ is excited then $g$ is also excited.\\
ii)  if the outputs are the same value and $f$ is unexcited then $g$ is also unexcited.\\
iii)  if the outputs are different values and $f$ is excited then $g$ is unexcited.\\
iv)  if $f$ is excited and $g$ is unexcited then the outputs have different values.
\end{lemma}
\begin{proof}
i) \& ii)  Since $f$ and $g$ are identical gates or C-elements then if the inputs and output are the same values, $f$ and $g$ must be both excited or both unexcited. \\
iii)  We prove this by contradiction.  Suppose that inputs to $f$ and $g$ are the same and the outputs are different values and both $f$ and $g$ are excited.  This means that for some circuit element that these input values both cause the element to excite to 1 and 0 at the same time.  This cannot happen in any of the elements that we build our circuit from and is therefore a contradiction.\\
iv)  We prove this by contradiction.  Suppose that inputs to $f$ and $g$ are the same and the outputs are the same values but $f$ is excited and $g$ is unexcited.  This violates i) and is therefore a contradiction.  
\end{proof}

@@

%P3 states that the wire block must be in a state a-d) of figure \ref{fig:l1helper}.  We consider what happens if a wire transitions from each state. 
%logic gate excited = c1^A excited
%possibly add graph to illustrate this
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{gatewl2}
  \caption[Different cases of a wireblock]{Different cases of P2 for a wireblock, 1-4 correspond to excited wire in the original circuit and 5 is unexcited.  1) case 1, 2) case 2, 3) case 3, 4) case 4, 5) case 5}
  \label{fig:l1helper}
\end{figure}

@@

\begin{figure}
  \centering
    \includegraphics{flowl2c3}
  \caption[Lemma 2 transitions]{The wires in each row are in order of logic block, C1, C2, C3, W.  Each row represents half of the duplicated circuit, ie. top row are A half wires and bottom row are B half wires.  'Un' denotes that the logic block is unexcited and 'Exc' denotes the logic block is excited.  Blue arrows indicate a case change caused by transitions elsewhere in the circuit.  Black arrows indicate a wire transition.}
  \label{fig:l21}
\end{figure}
%fixed 
\begin{lemma}[2.1]
For any trace of the fault-tolerant circuit, each wire block follow the state diagram in Figure \ref{fig:l21} (up to 0 and 1 inversion).
\end{lemma}
\begin{proof}
We show this lemma by induction.  The state diagram in Figure \ref{fig:l21} is incomplete in the sense that after case 11, there are more transitions.  However, case 11 is symmetric to case 1 with only the 0's and 1's inverted and so the transitions from case 11 is also symmetric to the transitions from case 1 with 0's and 1's inverted until it loops back to case 1.  Because of this symmetry, we only need to analyze possible transitions from each case up to case 10.  Also, Lemma 1 applies to both A half and B half which we will use in the body of this proof.\\

@@

%still a slight problem here
\end{itemize}
This shows that all wire blocks in all possible $s_{k+1}$ follow the state diagram, and the induction step is complete.\\
By induction on the consecutive states of the trace, the above argument shows that all wire blocks of a fault-tolerant circuit follow the state diagram in Figure \ref{fig:l21} for all traces. \\
\end{proof}
%may be that I don't need how all the intermediate wires go... but just need the at most one of input excitation and outputs can be mismatched.  But this needs to include if outputs are mismatched then the C-element is the same.... which I think I need to show the signals propagating through??
It is now easy to show that deadlock does not occur since each wire block must be in one of the configurations of Figure \ref{fig:l21}.
\begin{lemma}[2.2]
For all traces $\sigma$ in the fault-tolerant circuit, deadlock does not occur.
\end{lemma}
\begin{proof}
To show that deadlock does not occur, we show that we can always find an excited wire at any state $s_k$ in a trace of the fault-tolerant circuit.  The state $s_k$ can be split into two cases, one is that all output wire pairs in the A half and B half are equal, the second is that there exists at least one wire $v$ such that $v^A\neq v^B$.  In the first case, if every output wire pair are equal, then $h^A(s_n)=h^B(s_n)$ and there exists an excited wire $u$ in the original circuit.  By Lemma 1 both halves of the duplicated circuit for $u$ is excited.  By Lemma 2.1, the circuit must be in case 2-6 for $u$.  In each of these cases at least one wire is excited.  \\
In the second case, for some wire $v$ such that $v^A\neq v^B$, then using Lemma 2.1 wire block $v$ is in case 7a or 7b which has an excited wire, either $v^A$ or $v^B$.  Since there are next transitions for both cases, deadlock does not occur.
\end{proof}

Combining the results of the previous Lemmas shows that the fault-tolerant circuit is equivalent to the original circuit under normal operation.  This is defined formally in a theorem as follows:
\begin{theorem}[1]
For all traces $\sigma$ in the fault-tolerant circuit, $h^{A}(\sigma_n)$ or $h^{B}(\sigma_n)$ is a prefix of a trace in the original circuit for all $n \in \mathbb{N}$.  (h might make the trace shorter, so the prefix will not be of length n) %what was the point of the deadlock thing?
\end{theorem}
\begin{proof}
Apply Lemma 1 then $h^{A}(\sigma_n)$ or $h^{B}(\sigma_n)$ is a prefix of a trace.  From Lemma 2.2, no deadlock occurs so it is true for all $n\in \mathbb{N}$.
\end{proof}
%with these 3 properties show the original claim

\section{Single transient fault analysis}
%fix this
Next we consider what happens to the fault-tolerant circuit under a single transient fault.  A {\em transient fault} on a wire $w$ causes a transition i.e. flips the bit of $w$.  In light of this, we have expanded the definition of a transition to include the case where the wire value changes due to the transient fault.  After the transient event, the faulty wire transitions based on circuit rules.  %do I need to talk about when transient is not ~x?  ie hold same value?  That's the same as a longer delay
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{faultexes}
  \caption[Effects of a transient fault]{Effects of a transient fault.  The left most figure is the state right before a transient fault occurs.  The transition due to the transient fault is denoted by the red F and arrow.   \\
  a) An AND gate, the transient fault causes a transition in $w$ to 0 but the inputs resurrect it back to 1 sometime later.   \\
  b) An AND gate that is excited before the transient fault.  The transient fault causes a transition that would happen anyway, $w$ stays at the correct value 1 after the transition.\\
  c) A C-element.  Its inputs are mismatched so the gate is in a hold state.  When the transient fault causes a transition to 0, the faulty 0 value may persist.
  }
  \label{fig:faultx}
\end{figure}
Examples of the transient fault occuring on elements in different states are shown in Figure \ref{fig:faultx}.  Depending on the inputs and the element, the transient fault can cause a transition to a faulty value and transition back to the previous value; or it can cause a transition that would happen anyway and thus does not affect the element or circuit; or the faulty value may be stored in memory elements such as the C-element in the hold state.\\
